---
title: "Lab Assignment 3"
author: "Belina Jang"
date: "March 6, 2020"
output:
  pdf_document: default
  html_document: default
---

## General Instructions

* Put your name in the author section above.
* Write R code in the R chunks provided to answer the questions posed. Hint:  Refer to Lab6.Rmd.
* Execute each chunk of code to ensure that your code works properly.
* If you cannot compile all of your code without errors before the end of the class, 
comment out the chunks that are not working.
* Save this .Rmd file to the (Windows) C: drive so that there are no knitting issues and then knit the entire document to pdf.
* Copy your pdf and Rmd files to your personal directory (see the Lab 1 instructions if needed).
* Upload the pdf file to the Lab Assignment 3 Activity in the 
Lab section of CourseSpaces. 

## 1.  One of the most amazing results of probability is the Central Limit Theorem. [4 marks]
The Central Limit Theorem states that sums of independent, identically
distributed random variables from almost any probability distribution 
have distributions that can be approximately calculated with areas
under the normal distribution.

Repeat the investigation of the Central Limit Theorem
from Lab 6 using simulations of sums of exponential (mean=5) random variables.


```{r exp(1)}
set.seed(54321)
#Setting the random number generator seed SHOULD always be done before simulating
# Define a function to generate <repeats> replications of an experiment. 
#  For each replication, simulate sample <size> exponential random numbers.
#  Sum the <size> random numbers for each replication and return a vector of length <repeats>.
#  I code this efficiently without for loops using matrix computations instead of loops.
#  For loops are computationally expensive, and in R we try to avoid them when we can

Esim <- function(repeats, size){
  v <- rexp(size * repeats, rate = 5)   #generate size*repeats
  w <- matrix(v, nrow=size, ncol=repeats)  #Put v in a matrix with size rows and repeats cols
  return(colSums(w))  #Sum the columns, the result has length repeats
}

E1 <- Esim(10000, 1)   #size 1
head(E1)

E2 <- Esim(10000, 2)   #size 2
head(E2)

E5 <- Esim(10000, 5)   #size 5
head(E5)

E20 <- Esim(10000, 20)  #size 20
head(E20)     # sums of 20 E(0,1) random observations
length(E20)   # check the length, should be 10000

# Draw histograms of the 4 vectors.
#  look up the help file for the function layout
layout(matrix(c(1, 2, 3, 4), nrow=2, ncol=2, byrow=TRUE))  
hist(E1, prob=TRUE, ann=FALSE)   #ann=FALSE turns off annotation
title("Sample Size 1")

hist(E2, prob=TRUE, ann=FALSE) 
title("Sample Size 2")

hist(E5, prob=TRUE, ann=FALSE) 
title("Sample Size 5")

hist(E20, prob=TRUE, ann=FALSE) 
title("Sample Size 20")

layout(1)

```

## 2.  Repeat the investigation of the Central Limit Theorem using simulations of sums of Poisson (mean=1) random variables. [4 marks]

```{r poisson(5)}
set.seed(54321)   #Setting the random number generator seed SHOULD always be done before simulating
Esim <- function(repeats, size){
  v <- rexp(size * repeats, rate = 1)   #generate size*repeats
  w <- matrix(v, nrow=size, ncol=repeats)  #Put v in a matrix with size rows and repeats cols
  return(colSums(w))  #Sum the columns, the result has length repeats
}

E1 <- Esim(10000, 1)   #size 1
head(U1)

E2 <- Esim(10000, 2)   #size 2
head(U2)

E5 <- Esim(10000, 5)   #size 5
head(U5)

E20 <- Esim(10000, 20)  #size 20
head(E20)     # sums of 20 E(0,1) random observations
length(E20)   # check the length, should be 10000

# Draw histograms of the 4 vectors.
#  look up the help file for the function layout
layout(matrix(c(1, 2, 3, 4), nrow=2, ncol=2, byrow=TRUE))  
hist(E1, prob=TRUE, ann=FALSE)   #ann=FALSE turns off annotation
title("Sample Size 1")

hist(E2, prob=TRUE, ann=FALSE) 
title("Sample Size 2")

hist(E5, prob=TRUE, ann=FALSE) 
title("Sample Size 5")

hist(E20, prob=TRUE, ann=FALSE) 
title("Sample Size 20")

layout(1)

```


## 3. Using the Likelihood below, produce a three-dimensional graph of the Likelihood function surface for an independence model of section 4.6, where vector alpha and beta lie between 0 and 1. [5 marks]

The Likelihood function is:  alpha^76  (1-alpha)^24  beta^74  (1-beta)^26  


```{r 3Dim}
# Define the Likelihood function for vector alpha and vector beta.
#  Loop over the values for alpha and beta and create a matrix L.
#  L[i,j] contains the likelihood evaluated at alpha[i] and beta[j].

Like <- function(alpha,beta){
  L <- matrix(alpha^76  (1-alpha)^24  beta^74  (1-beta)^26)
  for(i in 1:{
    for(j in 1:{
      L[i,j] <- prod(dnorm(x, alpha[i], beta[j]))
    }
  }
  return(L)
  
  
  
}


# Create vector sequences, alphavals and betavals, which must lie between 0 and 1.
#  alphavals and betavals must contain the MLE, alphahat = .76 and betahat = .74

alphavals <-               # assign your values
betavals <-                # assign your values
  
# LL is a matrix of computed likelihood values for alphavals and betavals.

LL <- Like(alphavals, betavals)
  
  
# Produce a 3D perspective plot of the LL values.  Remember to title your plot.


```


