---
title: 'Assignment #4'
author: "Belina Jang"
date: "20/03/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
1. Student consultants at the HelpDesk encounter questions about programs written in Python,
Java, R and c#. The consultants have been hired based on the assumption that 40% of all
program questions concern Python, 25% concern Java, 25% concern R and 10% concern C#.
120 random program queries were tabulated and the resulting frequencies appear below.
Using a Likelihood ratio test and this data, test the null hypothesis that the probabilities for
the four programs are: .40, .25, .25 and .10 respectively. Comment on your findings including
a comparison of observed and expected frequencies under H0.

```{r question1}
freq <- c(52, 38, 21, 9)
sum(freq)
ell <- function(p, freq){
  # Multinomial log-likelihood
  # freq = frequencies;  p = probabilities
  sum(freq * log(p)) 
}

LRS <- function(p0, phat, freq){
  #Likelihood ratio statistic, phat=MLE, p0=H0 values
  2*(ell(phat, freq) - ell(p0, freq))
}

p0 <- c(0.40,0.25,0.25,0.10)
phat <- freq/sum(freq)
dobs <- LRS(p0, phat, freq)   #LRS observed
dobs
1 - pchisq(dobs, 3)
sum(freq)*p0 #expected frequencies
sum(freq)*phat #observed frequencies
```
The p-value for the null hypothesis is large (p-value:0.1053911 > 0.10), so there's no evidence against H0.

## Question 2

2. Three different regions off the coast of British Columbia are being considered as possible locations for a liquid natural gas (LNG) terminal. To get information on public opinion regarding
the LNG project, a separate random sample of individuals in each region is obtained, 400 from
region A and 350 each from regions B and C. Each individual is then asked whether he/she
favours building an LNG terminal at the proposed location in the region in which the individual
resides. The observed cell counts appear below.
Let pA, pB, pC be the true probabilities that individuals from the respective regions favour
the LNG project in their region. Using a Likelihood Ratio Test, test the null hypothesis that
pA = pB = pC. Comment on your findings including a comparison of observed and expected
frequencies of In Favour under H0.
Region A Region B Region C Total
In Favour 198 140 133 471
Against 202 210 217 629
Total 400 350 350 1100


```{r question2}
freq <- matrix(c(198,140,133,202,210,217), nrow=2, byrow=TRUE)
freq
sum(freq)

ell <- function(p, freq){
  # Multinomial log-likelihood
  # freq = frequencies;  p = probabilities
  sum(freq * log(p)) 
}

LRS <- function(p0, phat, freq){
  #Likelihood ratio statistic, phat=MLE, p=H0 values
  # freq=frequencies
  2*(ell(phat, freq) - ell(p0, freq))
}

rsum <- rowSums(freq)
csum <- colSums(freq)
rsum
csum

print("Expected Frequencies")
eij <- outer(rsum, csum) / sum(freq)  # expected freq - see help for outer
eij                                 # eij= r_i * c_j/n

# estimated probs under H0 are eij/sum(freq)
p0 <- eij / sum(freq)
phat <- freq / sum(freq)

dobs <- LRS(p0, phat, freq)
dobs
1 - pchisq(dobs, 2)  #pvalue, df=(#rows - 1)(#cols - 1)
```
The p-value 0.002851579 < 0.01, so there's strong evidence against H0.
And the observed frequencies for In favour under H0 was (198,140,133) looks bit different from its expected frequencies which was (171.2727 149.8636 149.8636).

## Question 3

3. A random sample of 1350 University students was asked about their political orientation and
their cannabis use. The table of frequencies appears below.
Using a Likelihood Ratio Test, test the null hypothesis that student political orientation is independent of their cannabis use. Comment on your findings including a comparison of observed
and expected frequencies under the null hypothesis of independence.
Political Never Use Cannabis Rarely Use Frequently Use
Liberal 479 173 119
Conservative 214 48 15
Other 172 45 85

```{r question3}
freq <- matrix(c(479,173,119,214,48,15,172,45,85), nrow=3, byrow=TRUE)
freq
sum(freq)

ell <- function(p, freq){
  # Multinomial log-likelihood
  # freq = frequencies;  p = probabilities
  sum(freq * log(p)) 
}

LRS <- function(p0, phat, freq){
  #Likelihood ratio statistic, phat=MLE, p=H0 values
  # freq=frequencies
  2*(ell(phat, freq) - ell(p0, freq))
}

rsum <- rowSums(freq)
csum <- colSums(freq)
rsum
csum

eij <- outer(rsum, csum) / sum(freq)  # expected freq - see help for outer
eij                                 # eij= r_i * c_j/n

# estimated probs under H0 are eij/sum(freq)
p0 <- eij / sum(freq)
phat <- freq / sum(freq)

dobs <- LRS(p0, phat, freq)
dobs
1 - pchisq(dobs, 4)  #pvalue, df=(#rows - 1)(#cols - 1)
```
```
The p-value is extremely small(<0.005) so we have very strong evidence H0. And the data suggest that there's an association between political orientation and cannabis use.
