---
title: "Lab Assignment 1"
author: "Belina Jang"
date: "January 31 2020"
output:
  pdf_document: default
  html_document: default
---

# Likelihood methods for the Poisson distribution

* Put your name in the author section above.
* Write R code in the R chunks provided to answer the questions posed.
* Execute each chunk of code to ensure that your code works properly.
* If you cannot compile all of your code without errors before the end of the class, 
comment out the chunks that are not working.
* Save this .Rmd file to the (Windows) C: drive so that there are no knitting issues and then knit the entire document to pdf.
* Copy your pdf and Rmd files to your personal directory (see the Lab 1 instructions if needed).
* Upload the pdf file to the Lab Assignment 1 Activity in the 
Lab section of CourseSpaces. 

## 1.  First generate 1 observation from the Poisson(lambda=5) distribution and print the value. [1 mark]


```{r data}
set.seed(12345)   #use this seed!
pdat1 <- rpois(1, lambda = 5)
pdat1
```

## 2.  Compute the Log-likelihood for a vector sequence of lambda values from 2 to 7 in steps of .05. 
(Hint:  See help for the R function called *dpois*.)

```{r LogLik}
#create a vector of values from 2 to 7 in steps of 0.05
lambda <- seq(2, 7, by=.05)

#compute the log of poisson probabilities
ploglike <- dpois(pdat1, lambda, log = TRUE)

#combine the lambda values and the log probabilities as column vectors
head(cbind(lambda, ploglike))
```


## 3. Plot the Log-Likelihood function values from your answer in 2 versus lambda. Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog}
lambda <- seq(2, 7, by=.05)
plot(ploglike ~ lambda, ylab='Log-Likelihood', xlab='lambda',type='b')
title('Poisson Log-likelihood function')
```


## 4.  Generate n=10 observations from the same distribution and compute and display the sample mean of the 10 observations.  (This is the MLE of lambda.) [1 mark]
(Hint:  See help for the R function called *mean*.)

```{r mean10}
set.seed(12345) #use this seed!
n <- 10
pdat2 <- rpois(n, 5)   # this generates n binomial observations
pdat2
mean(pdat2)
```

## 5.  Plot the log-likelihood function given the n=10 observations.  Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog10}

ploglike2 <- 0
for (i in 1:n){
  ploglike2 <- ploglike2 + dpois(pdat2[i],lambda, log=TRUE)  
  #computes the sum of the log of binomial probabilities over the observations
} 

head(cbind(lambda, ploglike2))

plot(ploglike2 ~ lambda, ylab='Log-Likelihood', xlab='lambda', type='b')
title(paste('Possion Log-likelihood for n=10'))
```

## 6.  Generate n=100 observations from the same distribution and compute and display the sample mean of the 100 observations.  (This is the MLE of lambda.) [1 mark]


```{r mean100}
set.seed(12345) #use this seed!
n <- 100
lambda <- seq(2, 7, by=.05)
pdat3 <- rpois(n, 5)   # this generates n binomial observations
pdat3
mean(pdat3)
```

## 7.  Plot the log-likelihood function given the n=100 observations.  Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog100}
ploglike3 <- 0
for (i in 1:n){
  ploglike3 <- ploglike3 + dpois(pdat3[i],lambda, log=TRUE)  
  #computes the sum of the log of binomial probabilities over the observations
} 

head(cbind(lambda, ploglike3))

plot(ploglike3 ~ lambda, ylab='Log-Likelihood', xlab='lambda', type='b')
title(paste('Possion Log-likelihood for n=100'))
```

